Understanding domestic abuse

### 2019-07-23 JeS_fields.tex

In 2018, there were X thousand victims of domestic abuse in England and Wales alone. Domestic abuse has a terrible impact upon victims in a variety of ways, and ultimately, the fundamental goal of this research project is to benefit victims of domestic abuse by improving their safety and access to support. However, the effects of domestic abuse are multi-faceted and far reaching. Such abuse often touches the lives of those far beyond the initial victim, such as children, family, neighbours, friends, and colleagues.

This research uses a variety of existing data sets, to help us better understand patterns of domestic abuse. The research will help to identify underlying risk factors for potential domestic abuse victims, and indicators for the most violent domestic abuse incidents such as domestic homicide. We will engage with a number of stakeholders, and provide insights relevant to their specific role in reducing the prevalence and harm caused by domestic abuse.

Many of our primary partners are within the police force. The research will help them to identify individuals who are most at risk of being victims of domestic abuse, and improve risk assessment models to identify situations which could escalate into serious acts of violence. It will also enable the identification of likely triggers for the onset, or worsening of ongoing abuse allowing police to better direct resources and design risk assessment frameworks. The potential triggers that we will investigate include high profile events such as sporting tournaments, and economic shocks, such as changes to the structure of benefit payments. Not only will this help the police, but it will also help policy makers estimate the likely impacts of policy change.

rates of gambling and alcohol spending in the local community, financial wellbeing, and

Ultimately, the fundamental goal of this research project is to benefit victims of domestic abuse by improving their safety and access to support. However, domestic abuse is multi-faceted and affects the lives of many individuals who  We aim to do so by working together other stakeholders, including the police, charities, and the government.

For charities and the government, insights from our research (risk factors of ) will provide useful information for the most effective targeting (geographical and temporal) of domestic abuse awareness campaigns.

 will benefit from a comprehensive investigation of a wide range of risk factors of domestic abuse victimisation.

Our insights about police mis-recording of domestic abuse will improve the experiences of victims in the CJS, increase their trust, and willingness to report future incidents. The police will also benefit from a deeper understanding of the factors governing reporting behaviour, its temporal characteristics throughout the year, and this knowledge will allow for an improved allocation of scarce police resources. Our research on the predictors of serious harm will help design better interventions.

victims of domestic abuse more liekly to report and less likely to be seriously harmed, children, police, policymaker - evidence base, charity

1) characteristics of victims and understanding predictors of serious harm - police can use insights to prevent serious harm, characteristics of victims - expand on previous qualitative research (academics) and help policy-makers better target campaigns

2) reporting behaviour and police mis-recording - police can improve recording practices, understanding reporting behaviour can benefit behavioural scientists who work for the government

3) consequences - helps the government and academics by advancing our knowledge on the consequences of domestic abuse victimisation by improving the methodology of

>>>>

In 2018, there were almost 2 million victims of domestic abuse in England and Wales alone. Domestic abuse has a terrible impact upon victims in a variety of ways, and ultimately, the fundamental goal of this research project is to benefit victims of domestic abuse by improving their safety and access to support. However, the effects of domestic abuse are multi-faceted and far reaching. Such abuse often touches the lives of those far beyond the initial victim, such as children, family, neighbours, friends, and colleagues.

This research uses a variety of existing data sets, to help us better understand the patterns, characteristics, and consequences of domestic abuse victimisation. The research will help to identify underlying risk factors for potential domestic abuse victims, and indicators for the most violent domestic abuse incidents, such as domestic homicide. We will engage with a number of stakeholders (police, domestic abuse charities, government), and provide insights relevant to their specific role in reducing the prevalence and harm caused by domestic abuse.

Many of our primary partners are within the police force. Our insights about police mis-recording of domestic abuse will help the police improve their recording practices, enhancing victim's trust in the police, and increase willingness to report future incidents.

In addition, the research will help the police to identify individuals who are most at risk of being victims of domestic abuse, and improve risk assessment models to identify situations which could escalate into serious acts of violence. It will also enable the identification of likely triggers for the onset, or worsening of ongoing abuse allowing police to better direct resources and design risk assessment frameworks. The potential triggers that we will investigate include high profile events such as sporting tournaments, and economic shocks, such as changes to the structure of benefit payments. Not only will this help the police, but it will also help policy makers estimate the likely impacts of policy change.

The police will also benefit from a deeper understanding of the factors governing reporting behaviour, and its temporal characteristics throughout the year, and this knowledge will allow them to improve the allocation of scarce police resources.

For domestic abuse charities and the government, insights from our research regarding the geographical and temporal characteristics of domestic abuse victimisation will provide invaluable information for the most effective targeting of domestic abuse awareness campaigns.
